Statistical Mechanics Assignment One

### 2016-04-15 stat_mech/latex/a1.tex
add new a1 statmech stuff.

>>>>

\section{Problem Four: Gas Engine}
Here we use the ideal gas property $dU = C_v$ to express all changes in internal energy
as changes in temperature. We also use the adiabatic $PV^\gamma = k$ relation to convert
between variables during the insulated work-doing steps:\\

We first find the work done:
\begin{align*}
  W_{ab} &= -dU \hspace{2cm}\mbox{Q = 0, First law of Thermo}\\
  &= -C_vdT \hspace{2cm}dU_{ideal} = CvdT\\
  &= -C_v\left(T_b-T_a\right)\\
  &= -C_vT_b\left(1-\frac{T_a}{T_b}\right)\\
  &= -C_vT_b\left(1-\left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
             \hspace{2cm}P_aV_2^\gamma = P_bV_1^\gamma \rightarrow
             \frac{T_a}{T_b} = \left(\frac{V_2}{V_1}\right)^{\gamma-1}\\
  W_{cd} &= C_vT_c\left(1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
\end{align*}

Then we find the heat added during the heating step:

\begin{align*}
  dQ_1 &= dU_V = C_v\left(T_c-T_b\right)\\
\end{align*}

Now we find the efficiency:

\begin{align*}
  \eta &= \frac{W}{Q_1} = \frac{C_vT_c\left(1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
    - C_vT_b\left(1-\left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)}{C_v\left(T_c-T_b\right)}\\
  &= 1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\\
\end{align*}
